### 2011-03-02 qualificacao/cap-ameacas.tex
TODOs introducao e espacos para resultados e ameacas

%% ------------------------------------------------------------------------- %%
\chapter{Ameaças a Validar}
\label{cap:ameacas}

Discutir ameaças a validar

\section{História de TDD}