### 2011-03-03 qualificacao/cap-conclusoes.tex
introducao

%% ------------------------------------------------------------------------- %%
\chapter{Conclusões}
\label{cap:conclusoes}

conclusoes

%------------------------------------------------------
\section{Sugestões para Pesquisas Futuras}