### 2011-07-18 qualificacao/ape-exercicios.tex
apendice de exercicios criado

\chapter{Exercícios}
\label{ape:exercicios}

\section{Exercício 1}

bla

\section{Exercício 2}

bla

\section{Exercício 3}

bla

\section{Exercício 4}

bla